validateHOT - Validate your Holdout Task

### 2023-11-11 README.Rmd
updated package - 231111 including markdown

[![license](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)
<!-- badges: end -->

The goal of `validateHOT` is to validate the results of your validation
task (also known as holdout task) as well as a tool for simulating markets with
the results of your MaxDiff, CBC, and ACBC. 
This package is especially relevant for the [Sawtooth Software](https://sawtoothsoftware.com/) community who would like to report their analysis in *R* for open science purposes. Of course,
this package is also useful for practitioners, who would like to run the analyses
in an open source software.
The ultimate goal of these preference measurement techniques is to predict 
future behavior (Green & Srinivasan, 1990). Therefore, it is essential for 
both academics and practitioners to ensure that the collected data is valid 
and can also predict outside tasks that were not included in the estimation of 
the utility scores.

@@

> `validateHOT` was developed to work with Sawtooth Software. Please be cautious about using it with different platforms (especially for linear and piecewise-coded variables).

@@

👈🏾 <u>**What you get**</u>:<br> At the moment, `validateHOT`
provides functions for 3 key areas:

@@

@@

For all 3 key areas, the `createHOT()` function is essential.
This function creates the total utilities for each alternative in the
validation/holdout task and also in the simulation method, respectively.
Rao (2014, p. 82) mentions the additive utility model stating that the
total utility of a profile or conjoint model is the sum of its attribute
levels. `createHOT()` will calculate the total utility of each alternative you specify.

> We are planning to expand the functions of the package by also providing functions to convert the raw utilities into scores that are easier to interpretate, for example, `prob_scores()` to calculate choice probabilities of both unanchored and anchored MaxDiff or `att_imp()` to calculate the importance of the attributes included in an ACBC or CBC.

@@

<li><code>hitrate()</code>: creates the Hit Rate (correctly predicted
choices) of your validation/holdout task. The output will contain 
the percentage of how many choices were correctly predicted, 
the chance level in your validation task ($\frac{1}{alternatives}$) in percentage,
the number of correctly predicted participants' choices as well as the number of observations. 
</li>

@@

<li><code>mae()</code>: calculates the mean absolute error, i.e., deviation between
predicted and stated choice share. </li>

@@

@@

<li><code>accuracy()</code>: calculates the number of correctly predicted choices (buy or no-buy);
$\frac{TP + TN}{TP + TN + FP + FN}$ (Burger, 2018). 
</li>

@@

<li><code>turf()</code>: <b>T</b>(otal) <b>U</b>(nduplicated) <b>R</b>(each) and <b>F</b>(requency) is a "product line extension model" (Miaoulis et al., 1990, p. 29) that helps to find the perfect product bundle based on the reach (e.g., how many participants consider buying at least one product of that assortment) and the frequency (how many products are on average a purchase option). <code>turf()</code> currently provides both the *threshold*  approach (<code>approach = 'thres'</code>; all products that exceed a threshold are considered, e.g., a purchase option; Chrzan & Orme, 2019, p. 112) and the *first choice* approach (<code>approach = 'thres'</code>; only product with highest utility is considered as purchase option; Chrzan & Orme, 2019, p. 111).</li>

@@

> <b>Important</b>: For both <code>reach()</code> and <code>freqassort()</code>, <code>noen</code> does not necessarily have to be the no-buy alternative but can be another alternative that should be exceeded.

@@

`validaeHOT` provides 5 data sets that come with the package. These datasets 
should help to better understand the way the data needs to be defined, 
especially for the `createHOT()` function.

@@

Now imagine you included a validation/holdout task with three
alternatives plus a no-buy alternative. We specify the `data` argument and the column index of `id`. Since we also have a *no-buy* alternative in our validation task, we next specify the `none` argument, otherwise we would have left it empty. Afterwards, we specify the number of alternatives (`prod`) in our validation/holdout task (excluding the *no-buy* alternative). Next, we define each alternative with the argument `prod.levels`. If we look back at the data frame, we can see that the first alternative in the holdout task (`c(4, 9, 19)`) is composed of the following attribute levels `Att1_Lev1`, `Att2_Lev1`, and `Att3_Lev5`.

@@

> &#10071; `createHOT()` is currently just taking indexes instead of column names, please be aware of this. However, you can easily find out the index by either using the `names()` function to see all names of the dataframe, or if you already know the name of the variable, you can use `which()` and `colnames()`, both functions are provided by the base package (R Core Team, 2023). For example, if you want to find out the column index of `ID`, you could run `which(colnames(CBC) == "ID")` to determine the column index of `ID`.

```{r}
HOT <- createHOT(
  data = CBC, # data frame
  id = 1, # column index of the id
  none = 21, # column index of none alternative
  prod = 3, # number of alternatives in validation task (excluding none)
  prod.levels = list(c(4, 9, 19), c(8, 12, 17), c(5, 10, 17)), # column index of the attribute levels for each attribute
  coding = c(0, 0, 0), # how the attributes were coded
  method = "CBC", # method
  choice = 22 # column index of choice
)
```

@@

In the next step, we would like to see how well our model (from which we took the raw utilities) predicts 
the actual choices in the validation/holdout task. First, we will run the `hitrate()` function. 
We specify the `data`, the column names of the alternatives (`opts`; remember 
there are three alternatives + the *no-buy* alternative), and finally the actual choice (`choice`).

@@

Next, we create the validation/holdout task to evaluate it in the next step. We use a validation/holdout task with three alternatives plus the *no-buy* alternative, just as we did in the previous example. The only difference to the previous example is that for our model estimation, the third attribute (`Att3_Lin`) was coded as linear.

Again, we first define data. The `id` is saved in the first column. The utilities for the `none` parameter are stored in the 15th column, and we again set `prod` to 3 (**excluding** the *no-buy* alternative). Next, we define the `prod.levels` for each alternative. For example, we can see that the second alternative is composed of `Att1_Lev5`, `Att2_Lev4`, and `60` which is the value that should be interpolated. We tell `createHOT()` that this value needs to be interpolated by specifying the `coding` argument. This tells us that the first two attributes (`Att1_Lev5`, `Att2_Lev4` in the case of alternative 2) are part-worth coded (`0`) while the third attribute is linear coded (`1`).

To interpolate the value, we have to provide `createHOT()` the `interpolate.levels`. These **need** to be the same as provided to Sawtooth as levels. Moreover, it is important that the value that should be interpolated needs to lie within the lower and upper bound of `interpolate.levels`. In our case, we had 7 levels that range from 10 to 70.

Next, we define the column index of the linear coded variable (`lin.p`) and specify the `coding` we talked about above. Again, we are running a *CBC* specified by the `method` argument. This time, we would like to keep some of the variables in the data frame, which we specify in `varskeep`. We only keep one further variable, however, you can specify as many as you want. This could be relevant if you would like to display results per group. Finally, we just tell `createHOT()` the column index of the final choice (`choice`) and we are all set.

```{r}
CBC <- createHOT(
  data = CBC_lin, # data frame
  id = 1, # column index of the id
  none = 15, # column index of none alternative
  prod = 3, # number of alternatives in validation task (excluding none)
  prod.levels = list(c(4, 9, 60), c(8, 12, 40), c(5, 10, 45)), # column index of the attribute levels for each attribute (for linear coded we specify value to be interpolated)
  interpolate.levels = list(c(10, 20, 30, 40, 50, 60, 70)), # actual values for the levels that should be interpolated
  lin.p = 14, # column index of the linear coded variable
  coding = c(0, 0, 1), # coding of the 3 attributes; 0 = part-worth, 1 = linear, 2 = piecewise
  method = "CBC", # method
  varskeep = 17, # column index of variables that should be kept in the dataframe
  choice = 16 # column index of choice
)
```

@@

Chrzan, K., & Orme, B. K. 2019. <em>Applied MaxDiff: A Practitioner’s Guide to Best-Worst Scaling</em> Provo, UT: Sawtooth Software.

Ding, Min, John R. Hauser, Songting Dong, Daria Dzyabura, Zhilin Yang, SU Chenting, and Steven P. Gaskin. 2011. “Unstructured Direct Elicitation of Decision Rules.” <em>Journal of Marketing Research 48</em>(1): 116-27. [https://doi.org/10.1509/jmkr.48.1.116](https://doi.org/10.1509/jmkr.48.1.116)

Drost, Hajk-Georg. 2018. “Philentropy: Information Theory and Distance Quantification with R” <em>Journal of Open Source Software 3</em>(26), 765, [https://joss.theoj.org/papers/10.21105/joss.00765](https://joss.theoj.org/papers/10.21105/joss.00765).

Gilbride, Timothy J., Lenk, Peter J., and Brazell, Jeff D. 2008. "Market Share Constraints and the Loss Function in Choice-Based Conjoint Analysis." <em>Marketing Science 27</em>(6):995-1011.[https://doi.org/10.1287/mksc.1080.0369](https://doi.org/10.1287/mksc.1080.0369)

Green, Paul E.,  and Srinivasan, V. 1990. "Conjoint Analysis in Marketing: New Developments with Implications for Research and Practice." <em>Journal of Marketing 54</em>(4): 3-19.[https://doi.org/10.1177/002224299005400402](https://doi.org/10.1177/002224299005400402).

Hamner, Ben, and Michael Frasco. 2018. “Metrics: Evaluation Metrics for Machine Learning.” [https://CRAN.R-project.org/package=Metrics](https://CRAN.R-project.org/package=Metrics).

@@

Orme, Bryan K. 2015. “Including Holdout Choice Tasks in Conjoint Studies.” [https://sawtoothsoftware.com/resources/technical-papers/including-holdout-choice-tasks-in-conjoint-studies](https://sawtoothsoftware.com/resources/technical-papers/including-holdout-choice-tasks-in-conjoint-studies).

Orme, B. K. (2020). <em>Getting Started with Conjoint Analysis: Strategies for Product Design and Pricing Research</em>. 4th edition. Manhattan Beach, CA: Research Publishers LLC.

R Core Team. 2023. “R: A Language and Environment for Statistical Computing.” [https://www.R-project.org/](https://www.R-project.org/).

Rao, Vithala R. 2014. <em>Applied Conjoint Analysis.</em> Springer Berlin Heidelberg. [https://doi.org/10.1007/978-3-540-87753-0](https://doi.org/10.1007/978-3-540-87753-0).

Yang, Liu (Cathy), Olivier Toubia, and Martijn G. de Jong. 2018. “Attention, Information Processing, and Choice in Incentive-Aligned Choice Experiments.” <em>Journal of Marketing Research 55</em>(6): 783–800. [https://doi.org/10.1177/0022243718817004](https://doi.org/10.1177/0022243718817004).

>>>>

[![license](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)

<!-- badges: end -->

The goal of `validateHOT` is to validate the results of your validation
task (also known as holdout task) as well as a tool for simulating
markets with the results of your MaxDiff, CBC, and ACBC. This package is
especially relevant for the [Sawtooth
Software](https://sawtoothsoftware.com/) community who would like to
report their analysis in *R* for open science purposes. However, it also
works perfectly with the great package `ChoiceModelR`(Sermas, 2022). Of
course, this package is also useful for practitioners, who would like to
run the analyses in an open source software. The ultimate goal of these
preference measurement techniques is to predict future behavior (Green &
Srinivasan, 1990). Therefore, it is essential for both academics and
practitioners to ensure that the collected data is valid and can also
predict outside tasks that were not included in the estimation of the
utility scores.

@@

> `validateHOT` was developed to work with Sawtooth Software and
> `ChoiceModelR`. Please be cautious about using it with different
> platforms (especially for linear and piecewise-coded variables).

@@

👈🏾 <u>**What you get**</u>:<br> At the moment, `validateHOT` provides
functions for 4 key areas:

@@

<li>converting raw logit utilities in scores that are easier to
interpret</li>

@@

For all 4 key areas, the `createHOT()` function is essential. This
function creates the total utilities for each alternative in the
validation/holdout task and also in the simulation method, respectively.
Rao (2014, p. 82) mentions the additive utility model stating that the
total utility of a profile or conjoint model is the sum of its attribute
levels. `createHOT()` will calculate the total utility of each
alternative you specify.

@@

<li><code>hitrate()</code>: creates the Hit Rate (correctly predicted
choices) of your validation/holdout task. The output will contain the
percentage of how many choices were correctly predicted, the chance
level in your validation task ($\frac{1}{alternatives}$) in percentage,
the number of correctly predicted participants' choices as well as the
number of observations.</li>

@@

<li><code>mae()</code>: calculates the mean absolute error, i.e.,
deviation between predicted and stated choice share.</li>

@@

Please be aware that for coding, `validateHOT` applies the following
coding of the *buy* and *no-buy* alternatives. Rows refer to the
observed decisions while columns refer to the predicted ones.

|        |     |        |
|--------|:---:|:------:|
|        | Buy | No-buy |
| Buy    | TP  |   FN   |
| No-buy | FP  |   TN   |

@@

<li><code>accuracy()</code>: calculates the number of correctly
predicted choices (buy or no-buy); $\frac{TP + TN}{TP + TN + FP + FN}$
(Burger, 2018).</li>

@@

<li><code>turf()</code>: <b>T</b>(otal) <b>U</b>(nduplicated)
<b>R</b>(each) and <b>F</b>(requency) is a "product line extension
model" (Miaoulis et al., 1990, p. 29) that helps to find the perfect
product bundle based on the reach (e.g., how many participants consider
buying at least one product of that assortment) and the frequency (how
many products are on average a purchase option). <code>turf()</code>
currently provides both the *threshold* approach (<code>approach =
'thres'</code>; all products that exceed a threshold are considered,
e.g., as purchase option; Chrzan & Orme, 2019, p. 112) and the *first
choice* approach (<code>approach = 'thres'</code>; only product with
highest utility is considered as purchase option; Chrzan & Orme, 2019,
p. 111).</li>

@@

> <b>Important</b>: For both <code>reach()</code> and
> <code>freqassort()</code>, <code>none</code> does not necessarily have
> to be the no-buy alternative but can be another alternative that
> should be exceeded.

### Reestimating raw utilities

<code>validateHOT</code> also provides 3 functions to better interpret
the scores of both (A)CBC and MaxDiff. Currently, 3 functions are
provided, namely:

<li><code>att_imp()</code>: converts the raw utilities of either an ACBC
or CBC into importance scores for each attribute (see, Orme, 2020, pp.
79-81)</li>

<li>

<code>prob_scores()</code>converts the raw utilities of a MaxDiff to
choice probabilities by applying the following procedures:

<li>For unanchored MaxDiff: First, the scores are zero-centered,
afterwards they are transformed by the following formula
$\frac{exp^{U_i}}{(exp^{U_i} + a - 1)}$ (Chrzan & Orme, 2019, p. 56),
where $U_i$ is the raw utility of item *i* and <code>a</code> is the
number of items shown simultaneously per MaxDiff task. Finally, the
scores are normalized.</li>

<li>For anchored MaxDiff:
$\frac{exp^{U_i}}{(exp^{U_i} + a - 1)} * 100 * \frac{1}{a}$ (Chrzan &
Orme, 2019, p. 59).</li>

</li>

<li><code>zc_diffs()</code>: rescales the raw logit utilities to make
them comparable across participants (Sawtooth Software Inc, 2023, p.
343).</li>

@@

`validaeHOT` provides 5 data sets that come with the package. These data
sets should help to better understand the way the data needs to be
defined, especially for the `createHOT()` function.

@@

Now imagine you included a validation/holdout task with 3 alternatives
plus a no-buy alternative. We specify the `data` argument and as well as
the `id`. Since we also have a *no-buy* alternative in our validation
task, we next specify the `none` argument, otherwise we would have left
it empty. Afterwards, we define each alternative with the argument
`prod.levels`. If we look back at the data frame, we can see that the
first alternative in the holdout task (`c(4, 9, 19)`) is composed of the
following attribute levels `Att1_Lev1`, `Att2_Lev1`, and `Att3_Lev5`.

@@

> ❗ `createHOT()` takes both column names and column indexes. However,
> please be aware, if you include linear coding or piecewise coding, you
> **have** to provide the column indexes for the input in prod.levels.,
> please be aware of this. However, you can easily find out the index by
> either using the `names()` function to see all names of the data
> frame, or if you already know the name of the variable, you can use
> `which()` and `colnames()`, both functions are provided by the base
> package (R Core Team, 2023). For example, if you want to find out the
> column index of `ID`, you could run `which(colnames(CBC) == "ID")` to
> determine the column index of `ID`.

```{r}
HOT <- createHOT(
  data = CBC,
  id = "ID",
  none = 21,
  prod.levels = list(c("Att1_Lev1", "Att2_Lev1", "Att3_Lev5"), c(8, 12, 17), c(5, 10, 17)),
  coding = c(0, 0, 0),
  method = "CBC",
  choice = "HOT"
)
```

@@

In the next step, we would like to see how well our model (from which we
took the raw utilities) predicts the actual choices in the
validation/holdout task. First, we will run the `hitrate()` function. We
specify the `data`, the column names of the alternatives (`opts`;
remember there are 3 alternatives + the *no-buy* alternative), and
finally the actual choice (`choice`).

@@

Next, we create the validation/holdout task to evaluate it in the next
step. We use a validation/holdout task with 3 alternatives plus the
*no-buy* alternative, just as we did in the previous example. The only
difference to the previous example is that for our model estimation, the
third attribute (`Att3_Lin`) was coded as linear.

Again, we first define data, namely the `id` as well as the `none`
alternative. Next, we define the `prod.levels` for each alternative.
Since we have one linear coded attribute, we have to specify the column
indexes instead of the column names. For example, for the second
alternative which is composed of `Att1_Lev5`, `Att2_Lev4`, and `60`
which is the value that should be interpolated. We tell `createHOT()`
that this value needs to be interpolated by specifying the `coding`
argument. This tells us that the first two attributes (`Att1_Lev5`,
`Att2_Lev4` in the case of alternative 2) are part-worth coded (`0`)
while the third attribute is linear coded (`1`).

To interpolate the value, we have to provide `createHOT()` the
`interpolate.levels`. These **need** to be the same as provided to
Sawtooth or `ChoiceModelR` as levels. Moreover, it is important that the
value that should be interpolated needs to lie within the lower and
upper bound of `interpolate.levels`. In our case, we had 7 levels that
range from 10 to 70.

Next, we define the column index of the linear coded variable (`lin.p`)
and specify the `coding` we talked about above. Again, we are running a
*CBC* specified by the `method` argument. This time, we would like to
keep some of the variables in the data frame, which we specify in
`varskeep`. We only keep one further variable, however, you can specify
as many as you want. This could be relevant if you would like to display
results per group. Finally, we define the actual choice (`choice`) in
the holdout/validation task and we are all set.

```{r}
CBC <- createHOT(
  data = CBC_lin, # data frame
  id = 1,
  none = 15,
  prod.levels = list(c(4, 9, 60), c(8, 12, 40), c(5, 10, 45)),
  coding = c(0, 0, 1),
  interpolate.levels = list(c(10, 20, 30, 40, 50, 60, 70)),
  lin.p = 14,
  method = "CBC",
  varskeep = "Group",
  choice = "HOT"
)
```

@@

Chrzan, K., & Orme, B. K. 2019. <em>Applied MaxDiff: A Practitioner's
Guide to Best-Worst Scaling</em> Provo, UT: Sawtooth Software.

Ding, Min, John R. Hauser, Songting Dong, Daria Dzyabura, Zhilin Yang,
SU Chenting, and Steven P. Gaskin. 2011. "Unstructured Direct
Elicitation of Decision Rules." <em>Journal of Marketing Research
48</em>(1): 116-27. <https://doi.org/10.1509/jmkr.48.1.116>

Drost, Hajk-Georg. 2018. "Philentropy: Information Theory and Distance
Quantification with R" <em>Journal of Open Source Software 3</em>(26),
765, <https://joss.theoj.org/papers/10.21105/joss.00765>.

Gilbride, Timothy J., Lenk, Peter J., and Brazell, Jeff D. 2008. "Market
Share Constraints and the Loss Function in Choice-Based Conjoint
Analysis." <em>Marketing Science
27</em>(6):995-1011.<https://doi.org/10.1287/mksc.1080.0369>

Green, Paul E., and Srinivasan, V. 1990. "Conjoint Analysis in
Marketing: New Developments with Implications for Research and
Practice." <em>Journal of Marketing 54</em>(4):
3-19.<https://doi.org/10.1177/002224299005400402>.

Sawtooth Software Inc. 2023. <em>Lighthouse Studio 9. Version
9.15.2.</em> Sequim, WA.: Sawtooth Software Inc.

Hamner, Ben, and Michael Frasco. 2018. "Metrics: Evaluation Metrics for
Machine Learning." <https://CRAN.R-project.org/package=Metrics>.

@@

Orme, Bryan K. 2015. "Including Holdout Choice Tasks in Conjoint
Studies."
<https://sawtoothsoftware.com/resources/technical-papers/including-holdout-choice-tasks-in-conjoint-studies>.

Orme, B. K. 2020. <em>Getting Started with Conjoint Analysis: Strategies
for Product Design and Pricing Research</em>. 4th edition. Manhattan
Beach, CA: Research Publishers LLC.

R Core Team. 2023. "R: A Language and Environment for Statistical
Computing." <https://www.R-project.org/>.

Rao, Vithala R. 2014. <em>Applied Conjoint Analysis.</em> Springer
Berlin Heidelberg. <https://doi.org/10.1007/978-3-540-87753-0>.

Sermas R (2022). *ChoiceModelR: Choice Modeling in R*. R package version
1.3.0, <https://CRAN.R-project.org/package=ChoiceModelR>.

Yang, Liu (Cathy), Olivier Toubia, and Martijn G. de Jong. 2018.
"Attention, Information Processing, and Choice in Incentive-Aligned
Choice Experiments." <em>Journal of Marketing Research 55</em>(6):
783--800. <https://doi.org/10.1177/0022243718817004>.

### 2023-11-11 paper/paper.Rmd
updated package - 231111 including markdown

validateHOT is a package that provides functions to both validate a validation/holdout task and run market simulations for results obtained in a (adaptive) choice-based conjoint analysis (hereafter ACBC and CBC, respectively) and maximum difference scaling (hereafter MaxDiff) using [Sawtooth Software](https://sawtoothsoftware.com/).[^1]

[^1]: validateHOT also works with MaxDiff raw logit utilities, see, for example, @chapman2023. For CBC, it should work with other software, as long as all attributes are part-worth coded. Linear and piecewise coding is implemented to work with Sawtooth Software.

Preference measurement techniques', such as (A)CBC or MaxDiff, ultimate goal is to predict future behavior [@green1990]. Hence, it is essential for both academics and practitioners to ensure that the collected data is valid and predicts outside tasks (i.e., the model has external validity) well. In terms of external validity, we refer to the generalizations to different settings [see, @calder1982, p.240]. The easiest way to test it is to include so-called validation or holdout tasks [@Orme2015], which are tasks that are fixed (i.e., same across participant) and are usually not used for estimating the part-worth utilities in hierarchical Bayes estimation. Practitioners often do not include them [@yang2018], which is unsatisfactory given the fact that the model is used to estimate market shares which poses the basis for relevant marketing decisions.

@@

validateHOT is a practical tool for Sawtooth Software users in industry as well as academia. It provides an open source solution for a) validating a validation/holdout task and ensuring that the model has predictive validity; b) running market simulations (e.g., **T**otal **U**nduplicated **R**each and **F**requency, hereafter TURF). Other packages, for example, Metrics [@Metrics] provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix. However, to put the Sawtooth export into the right format, the user needs some data wrangling which could pose a barrier. Moreover, there are also packages that however mainly focus on the analysis of conjoint analysis (e.g., ChoiceModelR [@ChoiceModelR], choicetools [@choicetools], logitR [@logitr], bayesm [@bayesm] etc.). To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations (especially TURF) is missing. validateHOT is especially helpful for researchers who plan to report results for research articles according to open science standards.

@@

validateHOT's functions can be categorized into four main components, see \autoref{tab:table1}. To bring the data into the right format, users can run the \texttt{\color{purple}createHOT} function, which creates the total utility of each alternative by applying the additive utility model [@rao2014, p. 82]. \colcod{turf} as well as the 2 rescaling functions, however, are not dependent on \texttt{\color{purple}createHOT}, and can be run using the raw logit scores.

| Validation metrics | Confusion matrix | Market simulations | Rescaling scores |
|:----------------:|:----------------:|:----------------:|:----------------:|
|     hitrate()      |    accuracy()    |    freqassort()    |    att_imp()     |
|        kl()        |       f1()       |     marksim()      |  prob_scores()   |
|       mae()        |   precision()    |      reach()       |                  |
|      medae()       |     recall()     |       turf()       |                  |
|       mhp()        |  specificity()   |                    |                  |
|       rmse()       |                  |                    |                  |

@@

To define the validation/holdout task, which has a total of 7 items (\texttt{\color{purple}prod}) plus the no-buy alternative (\texttt{\color{purple}none}), we use the \texttt{\color{purple}createHOT} function.

```{r}
# load data
data("MaxDiff")

# create holdout task and define relevant column indexes
HOT <- createHOT(
  data = MaxDiff, 
  id = 1, 
  none = 19, 
  prod = 7, # no of alternatives in HOT excluding none
  prod.levels = list(3, 10, 11, 15, 16, 17, 18), # index of alternatives
  method = "MaxDiff", 
  choice = 20, 
  varskeep = 21
)
```

@@

```{r}
# test hit rate of holdout task
hitrate(
  data = HOT, 
  opts = c(Option_1:None), 
  choice = choice 
) %>%
  round(3)
```

@@

```{r}
# test accuracy of holdout task
accuracy(
  data = HOT, 
  group = Group, 
  opts = c(Option_1:None), 
  choice = choice, 
  none = None 
)
```

Finally, we show two functions for market simulations, namely \texttt{\color{purple}marksim} and \texttt{\color{purple}turf}. In the following example, the market share is calculated accordning to the multinomial logit model [@McFadden1974].

```{r}
# run market simulations (share of preference rule)
marksim(
  data = HOT,
  opts = c(Option_1:None),
  method = "sop"
)
```

Finally, \texttt{\color{purple}turf}, a "product line extension model" [@miaoulis1990, p. 29], is a tool to find the perfect assortment that creates the highest reach and is especially powerful for MaxDiff studies [@chrzan2019, p. 108]. To optimize the search for the optimal bundle, we also include the arguments \texttt{\color{purple}fixed}, to define alternatives that have to be part of the assortment, and \texttt{\color{purple}prohib}, to prohibt certain item combinations of being part of the assortment (see the vignette for more details and how to apply \texttt{\color{purple}turf} with data obtained using a likert scale).

>>>>

validateHOT is a package that provides functions to both validate a validation/holdout task and run market simulations for results obtained in a (adaptive) choice-based conjoint analysis (hereafter ACBC and CBC, respectively) and maximum difference scaling (hereafter MaxDiff) using `ChoiceModelR` [@ChoiceModelR] or [Sawtooth Software](https://sawtoothsoftware.com/).

Preference measurement techniques', such as (A)CBC or MaxDiff, ultimate goal is to predict future behavior [@green1990]. Hence, it is essential for both academics and practitioners to ensure that the collected data is valid and predicts outside tasks (i.e., the model has external validity) well.[^1] The easiest way to test it is to include so-called validation or holdout tasks [@Orme2015], which are tasks that are fixed (i.e., same across participant) and are usually not used for estimating the part-worth utilities in hierarchical Bayes estimation. Practitioners often do not include them [@yang2018], which is unsatisfactory given the fact that the model is used to estimate market shares which poses the basis for relevant marketing decisions.

[^1]: In terms of external validity, we refer to the generalizations to different settings [see, @calder1982, p.240].

@@

validateHOT is a practical tool for Sawtooth Software users in industry as well as academia. It provides an open source solution for a) validating a validation/holdout task and ensuring that the model has predictive validity; b) running market simulations (e.g., **T**otal **U**nduplicated **R**each and **F**requency, hereafter TURF). Other packages, for example, Metrics [@Metrics] provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix. However, to put the Sawtooth export into the right format, the user needs some data wrangling which could pose a barrier. Moreover, there are also packages that however mainly focus on the analysis of conjoint analysis (e.g., ChoiceModelR [@ChoiceModelR], choicetools [@choicetools], logitR [@logitr], bayesm [@bayesm] etc.). To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations (especially TURF) is missing.

@@

validateHOT's functions can be categorized into four main components, see \autoref{tab:table1}. To bring the data into the right format, users can run the \texttt{\color{purple}createHOT} function, which creates the total utility of each alternative by applying the additive utility model [@rao2014, p. 82]. \colcod{turf} as well as the 3 rescaling functions, however, are not dependent on \texttt{\color{purple}createHOT}, and can be run using the raw logit scores.

| Validation metrics | Confusion matrix | Market simulations | Rescaling scores |
|:----------------:|:----------------:|:----------------:|:----------------:|
|     hitrate()      |    accuracy()    |    freqassort()    |    att_imp()     |
|        kl()        |       f1()       |     marksim()      |  prob_scores()   |
|       mae()        |   precision()    |      reach()       |    zc-diffs()    |
|      medae()       |     recall()     |       turf()       |                  |
|       mhp()        |  specificity()   |                    |                  |
|       rmse()       |                  |                    |                  |

@@

Assuming you included a validation/holdout task with a total of 7 alternatives plus the no-buy alternative (\texttt{\color{purple}none}). To create this validation task in *R*, we use the \texttt{\color{purple}createHOT} function.

```{r}
# create holdout task and define relevant column indexes
HOT <- createHOT(
  data = MaxDiff, 
  id = "ID", 
  none = "none", 
  prod.levels = list(3, 10, 11, 15, 16, 17, 18), 
  method = "MaxDiff", 
  choice = "HOT", 
  varskeep = "Group"
)
```

@@

```{r}
hitrate(
  data = HOT, 
  opts = c(Option_1:None), 
  choice = choice 
) %>%
  round(3)
```

@@

```{r}
accuracy(
  data = HOT, 
  group = Group, 
  opts = c(Option_1:None), 
  choice = choice, 
  none = None 
)
```

Finally, we show two functions for market simulations, namely \texttt{\color{purple}marksim} and \texttt{\color{purple}turf}. In the following example, the market share is calculated according to the multinomial logit model [@McFadden1974].

```{r}
marksim(
  data = HOT,
  opts = c(Option_1:None),
  method = "sop"
)
```

Finally, \texttt{\color{purple}turf}, a "product line extension model" [@miaoulis1990, p. 29], is a tool to find the perfect assortment that creates the highest reach and is especially powerful for MaxDiff studies [@chrzan2019, p. 108]. To optimize the search for the optimal bundle, we also include the arguments \texttt{\color{purple}fixed}, to define alternatives that have to be part of the assortment, and \texttt{\color{purple}prohib}, to prohibit certain item combinations of being part of the assortment (see the vignette for more details and how to apply \texttt{\color{purple}turf} with data obtained using a likert scale).
